### 2023-10-05 thesis/MSECE_thesis.tex
Thesis additions.

\section{Data Sources}\label{sec:data_source}
There are four datasets used in this work. The datasets are primarily speech corpora intended either for automatic speech recognition
research in the domain of aviation communications, linguistics research in aviation english, or both. As a result, the text-based data
samples that are used to train the language models in this work are transcripts of aural speech. The datasets used are described in more
detail below. All recordings and transcripts are in english, although there are multiple instances of foreign words occuring

@@

\subsection{ZCU CZ ATC Corpus}\label{sec:zcu_atc}
The ZCU CZ ATC corpus consists of audio recordings and corresponding transcripts in the Czech airspace. Both controller and pilot sides of the
communications are included in this data. Annotations were created by experienced transcribers/annotators and labeled samples were randomly selected
for review during the annoation process. After the dataset was completely labeled, all samples were checked, corrected, and
unified\cite{smidl_air_2019}.

@@

\subsection{Handling Spelling Errors \& Inconsistencies}\label{sec:spelling_errors}
By the nature of the data labeling process, there are bound to be spelling errors introduced by the annotators during corpora creation. This gets
exacerbated by the aggregation of several corpora with varying controls for correcting spelling errors and inconsistencies. Nearly all of the errors
are recoverable/correctable given context. For example, ``possibility of \textbf{tornado's}'' should clearly have been
``possibility of \textbf{tornadoes}'', changing the posessive form of ``tornado'' to the plural form. There are also instances of spelling conventions
that are consistent within corpora, but become inconsistent when the corpora are combined.

% 22 spelling errors
Manually reviewing every sample in the corpus is unrealistic considering that there are over 50,000 data samples and would be subject the same
human error that introduced those errors in the first place. Instead, the frequency of occurence of tokens in the corpora combined with manual review
thereafter, was used to detect and correct the most common errors. Tokens with the lowest frequency across corpora were manually reviewed and
corrected where necessary. An \(A \rightarrow B\) mapping system was used to correct the errors, where \(A\) is the erroneous token and \(B\) is the
corrected or replaced token. A total of 22 unique errors were detected and subsequently corrected or removed using this method.

% hesitations, cut off words, different kinds of filled pauses
The mapping system described above is also used to modify within corpora conventions to be consistent across corpora. Three areas are addressed here:
hesitations, cut-off or partial words, and filled pauses within speech. These areas are defined differently between the four corpora, which makes
unifying the transcripts difficult. Hesitations, for example, are defined by one corpus as any incomplete or non-english speech sounds and left
completely undefined by another (despite annotators making notes about hesitations). The way in which hesitations, pauses, and partial words are
transcribed differs between corpora as well. Two corpora prescribe special tokens to these phenomena such as ``[hes]'', ``[hesitation]'', ``<pause>'',
or similar, while the others transcribe the data as it sounds (i.e., ``uh'', ``eh'', or ``er'' for filled pauses and ``cir-'' or ``cir+'' for
instances that were cut-off mid pronunciation). In order to facilitate the unification of the transcripts for all four copora, all three of these
phenomena have been redefined with accompanying translation/replacement rules below:

\section{Data Analysis}\label{sec:data_analysis}
% sequence lengths, unique tokens, total number of tokens, total number of samples, breakdown by each dataset
Relevant corpora statistics were observed or calculated are are presented in Table~\ref{table:corpora_stats}. This includes the number of samples in
the corpus, mean sequence length per sample in the corpus, the number of unique tokens in the corpus, and the total number of tokens in the corpus.
The same statistics were also calculated across corpora, since all four corpora will be combined for training and testing the language and speech
recognition models. Note that this is based on the raw data after loading and unifying the corpora, prior to preprocessing for model training/testing
and outlier removal.

@@

The ATCC and ZCU CZ ATC corpora make up a significant portion of the data, ATCC alone makes up more than half (approximately 54\%) of the combined
data, and ATCC and ZCU CZ ATC together make up aprroximately 81\% of the data. The sequence lengths of the samples are fairly consistent across
corpora with at most about an 18\% percent difference between sequence lengths. The number of unique tokens does not scale additively to the
combined corpora, which is as expected. Considering that aviation English generally conforms to and follows ICAO phraseology standards we would expect
that many terms would be shared across corpora with the exception of geographic/location specific terms, such as city, landmark, airport, and airline
names, etc.~that typically vary across regions.

@@

\subsection{Data Preprocessing}
% explain and justify what is removed from the text, how, and why
% present and explain any relevant algorithms where necessary

\subsection{Tokenization}
% present and explain the tokenization algorithms used (possibly which models they correspond to)

>>>>

\section{Data Sources}\label{sec:data_source}
There are four datasets used in this work. The datasets are primarily speech corpora intended either for automatic speech recognition
research in the domain of aviation communications, linguistics research in aviation english, or both. As a result, the text-based data
samples that are used to train the language models in this work are transcripts of aural speech. The datasets used are described in more
detail below. All recordings and transcripts are in english, although there are multiple instances of foreign words occurring

@@

\subsection{ZCU CZ ATC Corpus}\label{sec:zcu_atc}
The ZCU CZ ATC corpus consists of audio recordings and corresponding transcripts in the Czech airspace. Both controller and pilot sides of the
communications are included in this data. Annotations were created by experienced transcribers/annotators and labeled samples were randomly selected
for review during the annotation process. After the dataset was completely labeled, all samples were checked, corrected, and
unified\cite{smidl_air_2019}.

@@

\subsection{Handling Spelling Errors \& Inconsistencies}\label{sec:spelling_errors}
By the nature of the data labeling process, there are bound to be spelling errors introduced by the annotators during corpora creation. This gets
exacerbated by the aggregation of several corpora with varying controls for correcting spelling errors and inconsistencies. Nearly all of the errors
are recoverable/correctable given context. For example, ``possibility of \textbf{tornado's}'' should clearly have been
``possibility of \textbf{tornadoes}'', changing the possessive form of ``tornado'' to the plural form. There are also instances of spelling conventions
that are consistent within corpora, but become inconsistent when the corpora are combined.

% 22 spelling errors
Manually reviewing every sample in the corpus is unrealistic considering that there are over 50,000 data samples and would be subject the same
human error that introduced those errors in the first place. Instead, the frequency of occurrence of tokens in the corpora combined with manual review
thereafter, was used to detect and correct the most common errors. Tokens with the lowest frequency across corpora were manually reviewed and
corrected where necessary. An \(A \rightarrow B\) mapping system was used to correct the errors, where \(A\) is the erroneous token and \(B\) is the
corrected or replaced token. A total of 22 unique errors were detected and subsequently corrected or removed using this method, although multiple
instances of those errors typically occurred.

% hesitations, cut off words, different kinds of filled pauses
The mapping system described above is also used to modify within corpora conventions to be consistent across corpora. Three areas are addressed here:
hesitations, cut-off or partial words, and filled pauses within speech. These areas are defined differently between the four corpora, which makes
unifying the transcripts difficult. Hesitations, for example, are defined by one corpus as any incomplete or non-english speech sounds and left
completely undefined by another (despite annotators making notes about hesitations). The way in which hesitations, pauses, and partial words are
transcribed differs between corpora as well. Two corpora prescribe special tokens to these phenomena such as ``[hes]'', ``[hesitation]'', ``<pause>'',
or similar, while the others transcribe the data as it sounds (i.e., ``uh'', ``eh'', or ``er'' for filled pauses and ``cir-'' or ``cir+'' for
instances that were cut-off mid pronunciation). In order to facilitate the unification of the transcripts for all four corpora, the conventions for
transcribing these phenomena are redefined below:

\begin{table}[h!]
    \centering
    \begin{tabular}{l p{0.65\linewidth}}
        \toprule
        Phenomenon               & Definition                                                                                                              \\
        \midrule
        Hesitations              & Instances of hesitation for which the specific realization is not defined (usually some special token is used instead). \\
        \midrule
        Partial \& cut-off words & Partially pronounced words are indicated by a hyphen character where the pronunciation of the word is missing.          \\
        \midrule
        Filled pauses            & Hesitations with a verbal component that has been described or otherwise defined by the corpus/annotator.               \\
        \bottomrule
    \end{tabular}
    \label{tab:phenomena_definitions}
    \caption{Speech/transcription phenomena with corresponding convention for transcribing or translating, depending on the corpus being processed.}
\end{table}

\begin{table}[h!]
    \centering
    \begin{tabular}{l l}
        \toprule
        Phenomenon                   & Convention for representation \\
        \midrule
        Hesitations                  & [HES]                         \\
        \midrule
        Partial \& cut-off words     & -                             \\
        \midrule
        Filled pauses (open-mouth)   & uh, um, ah                    \\
        \midrule
        Filled pauses (closed-mouth) & hm, mhm                       \\
        \bottomrule
    \end{tabular}
    \label{tab:phenomena_conventions}
    \caption{Conventions for transcript representation/translation of phenomena to keep transcripts consistent across corpora.}
\end{table}

A special token is used to represent hesitations when no other context regarding the realization of hesitation is present. If other corpora use a
special token to represent hesitations, it is translated to match the convention in table~\ref{tab:phenomena_conventions}. Partial/cut-off words
are represented with a hyphen (-) at the point where the pronounced word is cut off. For example, if the last syllable of ``approaching'' is cut
mid transmission, it would be represented in the transcript as ``approa-'' or, likewise, if the first syllable is cut off, it would be written as
``-roaching''. Lastly, if there is enough information in the original transcript to indicate the realization of a hesitation, it is represented by
the most appropriate token from ``Filled pauses'' in table~\ref{tab:phenomena_conventions}.

\section{Data Analysis}\label{sec:data_analysis}
% sequence lengths, unique tokens, total number of tokens, total number of samples, breakdown by each dataset
Relevant corpora statistics were observed or calculated and are presented in Table~\ref{table:corpora_stats}. This includes the number of samples in
the corpus, mean sequence length per sample in the corpus, the number of unique tokens in the corpus, and the total number of tokens in the corpus.
The same statistics were also calculated across corpora, since all four corpora will be combined for training and testing the language and speech
recognition models. Note that this is based on the raw data after loading and unifying the corpora, prior to preprocessing for model training/testing
and outlier removal.

@@

The ATCC and ZCU CZ ATC corpora make up a significant portion of the data, ATCC alone makes up more than half (approximately 54\%) of the combined
data, and ATCC and ZCU CZ ATC together make up approximately 81\% of the data. The sequence lengths of the samples are fairly consistent across
corpora with at most about an 18\% percent difference between sequence lengths. The number of unique tokens does not scale additively to the
combined corpora, which is as expected. Considering that aviation English generally conforms to and follows ICAO phraseology standards we would expect
that many terms would be shared across corpora with the exception of geographic/location specific terms, such as city, landmark, airport, and airline
names, etc.~that typically vary across regions.

@@

\section{Language Modeling}
% bert, roberta
% ngram
This section introduces the chosen language model architectures as well as their pretrained checkpoints, if one is used. Two transformer-based
architectures are used, BERT and RoBERTa, and statistical/machine learning-based approaches such as N-gram and word lattices.

The transformers are pre-trained using the same masked language modeling (MLM) objective presented in Devlin et al.~\cite{devlin_bert_2019}
to train BERT for bidirectional encoding representation. Both pretrained checkpoints and base models are pretrained and used with the best performing
ASR model to analyze their impacts on performance. The pre-training data is taken from the unified transcripts as described above, tokens are masked
at random with a probability of 0.2.

\section{Speech Recognition}

\section{Tokenization Algorithms}
% present and explain the tokenization algorithms used (possibly which models they correspond to)
Tokenization is the process of separating or segmenting samples into tokens. The rules by which the tokens are created or segmented depends entirely
on the algorithm used. The simplest and most intuitive way to do this by creating tokens based on whitespace i.e., each ``word'' in the corpus
is treated as a token\footnote{The Huggingface Tokenizer library implements this as the \lstinline|WordLevel| tokenizer.}. Although this is a simple
and easy to understand method of tokenization, it does little to nothing to address the out-of-vocabulary problem that necessarily comes with any
NLP tasks~\cite{wu_googles_2016}.

\subsection{WordPiece}

\subsection{(Byte-Level) Byte-Pair Encoding}
